Text embedding models enable high-resolution insights into conceptual knowledge and learning in classroom-like settings

### 2023-01-31 paper/main.tex

We used our model to transform the text of each question in our pool into
the same embedding space as the two lectures' trajectories 
(Fig.~\ref{fig:sliding-windows}C). This yielded a single 15-dimensional
coordinate (i.e., topic vector) for each question. We then computed the variance
of each topic's weight across timepoints of each lecture, and across 
questions from each category (i.e., lecture 1-related, lecture 2-related, and
general physics knowledge; Fig.~\ref{fig:topics}A). Visual inspection of 
Figure~\ref{fig:topics}A reveals a strong correspondence between the sets of 
topics expressed by each lecture and its related questions, as well as limited 
overlap in the topics expressed by non-matched lecture-question set pairs. 
To quantify these apparent similarities and differences, we computed the correlation 
between each pair of topic-weight variance distributions (Fig.~\ref{fig:topics}B). 
We found that our model captured the contents of lecture-related questions 
using a weighted blend of topic dimensions that was highly similar to that of 
their reference lecture (\textit{Four Fundamental Forces} (\textit{FFF}) 
questions vs.~lecture: Pearson's $r(13) = 
0.824,~p<0.001,~95\%~\mathrm{confidence~interval~(CI)} = [0.696, 0.973]$; 
\textit{Birth of Stars} (\textit{BoS}) questions vs.~lecture: $r(13) = 
0.801,~p<0.001,~95\%~\mathrm{CI} = [0.539, 0.958]$) but diverged significantly 
from that of the non-reference lecture (\textit{FFF} questions vs.~\textit{BoS} 
lecture: $r(13) = -0.620,~p = 0.014,~95\%~\mathrm{CI} = [-0.871, -0.326]$; 
\textit{BoS} questions vs.~\textit{FFF} lecture: $r(13) = -0.550,~p = 
0.034,~95\%~\mathrm{CI} = [-0.803, -0.246]$). This indicated that our model 
was sufficiently sensitive to the conceptual content tested by different quiz questions to differentiate between those relating to one lecture versus the other, and allow us to

sensitive to the underlying conceptual 
content tested by different question,and should allow us to differentially assess 
participants' knowledge for the two lectures based on the \textit{specific}
questions they answered correctly or incorrectly on each quiz.
%When we compared the
%variability in topic weights across each lecture's timepoints, and across
%questions from each category (i.e., lecture 1-related, lecture 2-related, and
%general physics knowledge), we found multiple interesting patterns (Fig.~\ref{fig:topics}). First,
%there is a strong correspondence between the sets of topics expressed by each
%lecture and its relevant questions. This indicates that our model is
%sensitive to the underlying conceptual content tested by different question,
%and should allow us to differentially assess participants' knowledge for the
%two lectures based on the \textit{specific} questions they answered correctly
%or incorrectly on each quiz. Second, the topics expressed by the general
%physics knowledge questions comprise a nonspecific mix of topics from the two
%lectures (4 from lecture 1; 5 from lecture 2), most of which are expressed to
%a relatively low degree. This indicates that our model should enable us to
%discriminate questions that are highly relevant to our assessment of
%participants' knowledge about concepts from the lectures from those that are
%not (or potentially less so). Third, the two lectures' trajectories vary
%along largely non-overlapping topic dimensions (e.g., variations in weight
%for topics 2, 3, 8, 12, 13, and 15 are specific to lecture 1; variations in
%weight for topics 1, 4, 5, 6, 7, and 10 are specific to lecture 2), but also
%share a small number of expressed topics in common (topics 11 and 14). This
%indicates that our lecture is able to simultaneously capture

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{figs/lecture-question-similarity}
    \caption{\textbf{Which parts of each lecture are captured by each
    question?} Each panel displays timeseries plots showing how each question's
    topic vector correlates with each video timepoint's topic vector (Panel
    \textbf{A.}: correlations for the \textit{Four Fundamental Forces} lecture
    and associated questions; Panel \textbf{B.}: correlations for the
    \textit{Birth of Stars} lecture and associated questions). The colors denote
    question identities. The diamonds in each panel denote the moment of peak
    correlation between the indicated questions, in the indicated lectures. The
    associated questions' text, and snippets of the lectures' transcripts in
    the best-matching sliding windows, are displayed at the bottom of the
    figure.}

@@

    \caption{\textbf{Mapping out the geometry of knowledge and learning.}
    \textbf{A. Average ``knowledge maps'' estimated using each quiz.} Each map
    displays a 2D projection of the estimated knowledge about the content
    reflected by \textit{all} regions of topic space (see
    \nameref{subsec:knowledge-maps}). The topic trajectories of each
    lecture and the coordinates of each question are indicated by dotted lines
    and dots. Each map reflects an average across all participants. For
    individual participants' maps, see
    Figures~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,
    and~\individualKnowledgeMapsC. \textbf{B. Average ``learning maps''
    estimated between each successive pair of quizzes.} The learning maps are
    in the same general format as the knowledge maps in Panel A, but each
    coordinate in the learning maps indicates the \textit{difference} between
    the corresponding coordinates in the indicated \textit{pair} of knowledge
    maps---i.e., how much the estimated knowledge ``changed'' across the two
    quizzes. Each map reflects an average across all participants. For
    individual participants' maps, see
    Figures~\individualLearningMapsA~and~\individualLearningMapsB. \textbf{C.
    Word clouds for sampled points in topic space.} Each word cloud displays
    the relative weights of each word reflected by the blend of topics
    represented at the locations of the stars in the maps. The words' colors
    indicate how much each word is weighted on average across all timepoints'
    topic vectors in the \textit{Four Fundamental Forces} (blue) and
    \textit{Birth of Stars} (green) videos, respectively.}

@@

We projected the resulting 100-dimensional topic vectors (for each second of
video and for each question) into a shared 2-dimensional space (see
\nameref{subsec:knowledge-maps}). Next, we sampled
points evenly from a $100 \times 100$ grid of coordinates that evenly tiled a
rectangle enclosing the 2D projections of the videos and questions. We used
Equation~\ref{eqn:rbf-knowledge} to estimate participants' knowledge at each of
these 10K sampled locations, and we averaged these estimates across
participants to obtain an estimated average \textit{knowledge map}
(Fig.~\ref{fig:knowledge-maps}). Intuitively, the knowledge map constructed
from a given quiz's responses provides a visualization of how ``much''
participants know about any content expressible by the fitted text embedding
model.

@@

Because the 2D projection we used to construct the knowledge and learning maps
is (partially) invertible, we may gain additional insights into the estimates
by reconstructing the original high-dimensional topic vectors for any point(s)
in the maps we are interested in. For example, this could serve as a useful
tool for an instructor looking to better understand which content areas a
student (or a group of students) knows well (or poorly). As a demonstration, we
show the top-weighted words from the blends of topics reconstructed from three
example locations on the maps (Fig.~\ref{fig:knowledge-maps}C): one point near
the \textit{Four Fundamental Forces} embedding (yellow); a second point near
the \textit{Birth of Stars} embedding (orange), and a third point somewhere in
between the two lectures' embeddings (pink). As shown in the word clouds in the
Panel, the top-weighted words at the example coordinate near the \textit{Four
Fundamental Forces} embedding also tended to be weighted heavily by the topics
expressed in that lecture. Similarly, the top-weighted words at the example
coordinate near the \textit{Birth of Stars} embedding tended to be weighted
most heavily by the topics expressed in \textit{that} lecture. And the
top-weighted words at the example coordinate between the two lectures'
embeddings show a roughly even mix of words most strongly associated with each
lecture.

@@

Over the past several years, the global pandemic has forced many educators to
teach remotely~\citep{MoseEtal21, ShimLee20, KawaEtal21, Whal20}. This change
in world circumstances is happening alongside (and perhaps accelerating)
geometric growth in the availability of high quality online courses on
platforms such as Khan Academy~\citep{Khan04}, Coursera~\citep{Youn12},
EdX~\citep{Kolo13}, and others~\citep{RhoaEtal13}. Continued expansion of the
global internet backbone and improvements in computing hardware have also
facilitated improvements in video streaming, enabling videos to be easily
downloaded and shared by large segments of the world's population. This
exciting time for online course instruction provides an opportunity to
re-evaluate how we, as a global community, educate ourselves and each other.
For example, we can ask: what makes an effective course or training program?
Which aspects of teaching might be optimized or automated? How and why do
learning needs and goals vary across people? How might we lower barriers to
achieving a high quality education?

@@

\subsubsection*{Constructing text embeddings of multiple lectures and questions}\label{subsec:topic-modeling}
Following~\citet{HeusEtal21}

We adapted an approach we developed in prior work~\citep{HeusEtal21} to embed
each moment of the two lectures and each question in our pool in a common representational space.
Briefly, our approach uses a topic model~\citep[Latent Dirichlet Allocation;][]{BleiEtal03},
fit to a corpus of documents to discover a set of $k$ ``topics'' or ``themes'' from their contents.
Formally, each topic is defined as a set of weights over each word in the model's vocabulary
(i.e., the union of all unique words, across all documents, excluding ``stop
words.'').

We adapted an approach we developed in prior work~\citep{HeusEtal21} to embed
each moment of the two lectures and each question in our pool in a common representational space.
Briefly, our approach uses a topic model~\citep[Latent Dirichlet Allocation;][]{BleiEtal03}, trained on a set
of documents, to discover a set of $k$ ``topics'' or ``themes.'' Formally, each
topic is defined as a set of weights over each word in the model's vocabulary
(i.e., the union of all unique words, across all documents, excluding ``stop
words.''). Conceptually, each topic is intended to give larger weights to words
that are conceptually related or that tend to co-occur in the same documents.
After fitting a topic model, each document in the training set, or any
\textit{new} document that contains at least some of the words in the model's
vocabulary, may be represented as a $k$-dimensional vector describing how much
the document (most probably) reflects each topic. (Unless, otherwise noted, we
used $k = 15$ topics.)

%\begin{figure}[tp]
%\centering
%\includegraphics[width=\textwidth]{figs/sliding_windows}
%
%\caption{\textbf{Constructing video content \textit{trajectories}. \textbf{A}.
%Building a document pool from sliding windows of text.} We decompose each
%video's transcript into a series of overlapping sliding windows. The set of
%transcript snippets (across all windows) may be treated as a set of
%``documents'' for training a text embedding model. After training a text
%embedding model on the two videos' sliding windows, we construct ``trajectories''
%through text embedding space by joining the embedding coordinates of successive
%sliding windows from each video. \textbf{B. Embedding multiple videos and
%questions.} Applying the same text embedding approach to each video, along with
%the text of each question in our pool (Tab.~\questions), results in one trajectory per video and one
%embedding coordinate (dot) per question (blue: \textit{Four Fundamental
%Forces}; green: \textit{Birth of Stars}; purple: general physics knowledge). \color{red}{Here
%we have projected the 15-dimensional embeddings into a 3D space using Uniform
%Manifold Approximation and Projection~\citep[UMAP;][]{McInEtal18a}}.}
%
%\label{fig:sliding-windows}
%\end{figure}

As illustrated in Figure~\ref{fig:sliding-windows}A, we first

As illustrated in Figure~\ref{fig:sliding-windows}A, we start by building up a
corpus of documents using overlapping sliding windows that span each video's
transcript. Khan Academy provides professionally created, manual transcriptions of all videos for closed captioning. However, such transcripts would not be readily available in all contexts to which our framework could potentially be applied. Khan Academy videos are hosted on the YouTube platform, which additionally provides automated transcripts

Khan Academy videos are hosted on the YouTube platform, and all
YouTube videos are run through Google's speech-to-text API~\citep{HalpEtal16}
to derive a timestamped transcript of any detected speech in the video. The
resulting transcripts contain one timestamped row per line, and each line
generally corresponds to a few seconds of spoken content from the video. We
defined a sliding window length of (up to) $w = 30$ transcript lines, and we
assigned each window a timestamp according to the midpoint between its first
and last lines' timestamps. These sliding windows ramped up and down in length
at the very beginning and end of the transcript, respectively. In other words,
the first sliding window covered only the first line from the transcript; the
second sliding window covered the first two lines; and so on. This insured that
each line of the transcript appeared in the same number ($w$) of sliding
windows. We treated the text from each sliding window as a single ``document,''
and we combined these documents across the two videos' windows to create a single
training corpus for the topic model.  The top words from each of the 15 discovered
topics may be found in Table~\topics.

After fitting a topic model to each videos' transcripts, we could use the
trained model to transform arbitrary (potentially new) documents into
$k$-dimensional topic vectors. A convenient property of these topic vectors is
that documents that reflect similar blends of topics (i.e., documents that
reflect similar themes, according to the model) will yield similar (in terms of
Euclidean distance, correlation, etc.) topic vectors. In general, the
similarity between different documents' topic vectors may be used to
characterize the similarity in conceptual content between the documents.

We transformed each sliding window's text into a topic vector, and then used
linear interpolation (independently for each topic dimension) to resample the
resulting timeseries to once per second. This yielded a single topic vector for
each second of each video. We also used the fitted model to obtain topic
vectors for each question in our pool (Tab.~\questions). Taken together, we
obtained a \textit{trajectory} for each video, describing its path through
topic space, and a single coordinate for each question
(Fig.~\ref{fig:sliding-windows}B). Embedding both videos and all of the
questions using a common model enables us to compare the content from different
moments of videos, compare the content across videos, and estimate potential
associations between specific questions and specific moments of video.

@@

\begin{equation}
    \mathrm{ncorr}(x, y) = \frac{\mathrm{corr}(x, y) - \mathrm{mincorr}}{\mathrm{maxcorr} - \mathrm{mincorr}},
\end{equation}
and where $\mathrm{mincorr}$ and $\mathrm{maxcorr}$ are the minimum and maximum
correlations between any lecture timepoint and question, taken over all
timepoints and questions across both lectures and all three question sets. We
also define $f(s, \Omega)$ as the $s$\textsuperscript{th} topic vector from the
set of topic vectors $\Omega$. Here $t$ indexes the set of lecture topic
vectors, $L$, and $i$ and $j$ index the topic vectors of questions in the
quiz's question set, $Q$. Note that ``correct'' denotes the set of indices of
the questions the participant answered correctly on the given quiz.

Intuitively, $\mathrm{ncorr}(x, y)$ is the correlation between two topic
vectors (e.g., the topic vector from one timepoint in a lecture, $x$, and the
topic vector for one question, $y$), normalized by the minimum and maximum
correlations (across all timepoints and questions) to range between 0 and 1,
inclusive. Equation~\ref{eqn:prop} then computes the weighted average
proportion of correctly answered questions about the content presented at
timepoint $t$, where the weights are given by the normalized correlations
between timepoint $t$'s topic vector and the topic vectors for each question.
The normalization step (i.e., using $\mathrm{ncorr}$ instead of the raw
correlations) insures that every question (except the least-relevant question)
contributes some non-zero amount to the knowledge estimate.

@@

An important feature of our approach is that, given a trained text embedding
model and participants' quiz performance on each question, we can estimate
their knowledge about \textit{any} content expressible by the embedding model--
not solely the content explicitly probed by the quiz questions. To visualize
these estimates
(Figs.~\ref{fig:knowledge-maps},~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,~\individualKnowledgeMapsC,~\individualLearningMapsA,
and~\individualLearningMapsB), we used UMAP~\citep{McInEtal18a} to define a 2D
projection of the text embedding space. Sampling the original 100-dimensional
space at high resolution to obtain an adequate set of topic vectors spanning
the embedding space would be computationally intractable. However, sampling a
2D grid is much more feasible. We defined a rectangle enclosing the 2D
projections of the lectures' and quizzes' embeddings, and we sampled points
from a regular $100 \times 100$ grid of coordinates that evenly tiled the
enclosing rectangle. We sought to estimate participants' knowledge (and
learning---i.e., changes in knowledge) at each of the resulting 10000
coordinates.

>>>>

We used our model to transform the text of each question in our pool into
the same embedding space as the two lectures' trajectories 
(Fig.~\ref{fig:sliding-windows}C). This yielded a single 15-dimensional
coordinate (i.e., topic vector) for each question. We then computed the variance
of each topic's weight across timepoints of each lecture, and across 
questions from each category (i.e., lecture 1-related, lecture 2-related, and
general physics knowledge; Fig.~\ref{fig:topics}A). Visual inspection of 
Figure~\ref{fig:topics}A suggests a strong correspondence between the sets of 
topics expressed by each lecture and its related questions, with only limited 
overlap in the topics expressed by non-matched lecture-question set pairs. 
To quantify these apparent similarities and differences, we computed the correlation 
between each pair of topic-weight variance distributions (Fig.~\ref{fig:topics}B). 
We found that our model represented the contents of lecture-related questions 
using a combination of topic dimensions (and relative variation along those topic dimensions) that was highly similar to that of 
their reference lecture (\textit{Four Fundamental Forces} (\textit{FFF}) 
questions vs.~lecture: Pearson's $r(13) = 
0.824,~p<0.001,~95\%~\mathrm{confidence~interval~(CI)} = [0.696, 0.973]$; 
\textit{Birth of Stars} (\textit{BoS}) questions vs.~lecture: $r(13) = 
0.801,~p<0.001,~95\%~\mathrm{CI} = [0.539, 0.958]$) but diverged significantly 
from that of the non-reference lecture (\textit{FFF} questions vs.~\textit{BoS} 
lecture: $r(13) = -0.620,~p = 0.014,~95\%~\mathrm{CI} = [-0.871, -0.326]$; 
\textit{BoS} questions vs.~\textit{FFF} lecture: $r(13) = -0.550,~p = 
0.034,~95\%~\mathrm{CI} = [-0.803, -0.246]$). This indicated that our model
captured the conceptual contents of the lectures and quiz questions 
sufficiently well to differentiate between questions relating to one lecture 
versus the other.

%This indicated that our model
%captures the conceptual contents of the lectures and quiz questions 
%sufficiently well to differentiate between questions relating to one lecture 
%versus the other, and could potentially be used to assess participants' knowledge for each lecture

% , and could be used to assess participants' knowledge about
%each lecture independently, based on the \textit{specific} questions they
%answered correctly or incorrectly on each quiz. 
%providing a slightly more 
%nuanced measure of what concepts they do and do not know than a simple 
%proportion-correct score.

%When we compared the
%variability in topic weights across each lecture's timepoints, and across
%questions from each category (i.e., lecture 1-related, lecture 2-related, and
%general physics knowledge), we found multiple interesting patterns (Fig.~\ref{fig:topics}). First,
%there is a strong correspondence between the sets of topics expressed by each
%lecture and its relevant questions. This indicates that our model is
%sensitive to the underlying conceptual content tested by different question,
%and should allow us to differentially assess participants' knowledge for the
%two lectures based on the \textit{specific} questions they answered correctly
%or incorrectly on each quiz. Second, the topics expressed by the general
%physics knowledge questions comprise a nonspecific mix of topics from the two
%lectures (4 from lecture 1; 5 from lecture 2), most of which are expressed to
%a relatively low degree. This indicates that our model should enable us to
%discriminate questions that are highly relevant to our assessment of
%participants' knowledge about concepts from the lectures from those that are
%not (or potentially less so). Third, the two lectures' trajectories vary
%along largely non-overlapping topic dimensions (e.g., variations in weight
%for topics 2, 3, 8, 12, 13, and 15 are specific to lecture 1; variations in
%weight for topics 1, 4, 5, 6, 7, and 10 are specific to lecture 2), but also
%share a small number of expressed topics in common (topics 11 and 14). This
%indicates that our lecture is able to simultaneously capture

This could enable us to estimate participants' knowledge for each of the two lectures, separately, by

This aspect of our model could enable us to assess participants' knowledge for each lecture separately, based on which \textit{specific} questions they answered right on wrong on each quiz, providing a slightly more refined estimate of the concepts they do and do not know than single proportion-correct score. However, while lectures are often organized around a single, broad theme at a coarse timescale, they typically cover...
%While individual lectures may be organized around a single broad theme at a coarse scale,

\textcolor{red}{\textbf{Paxton stopped here}}

\begin{figure}[tp]
    \centering
    \includegraphics[width=\textwidth]{figs/lecture-question-similarity}
    \caption{\textbf{Which parts of each lecture are captured by each
    question?} Each panel displays timeseries plots showing how each question's
    topic vector correlates with each video timepoint's topic vector (Panel
    \textbf{A.}: correlations for the \textit{Four Fundamental Forces} lecture
    and associated questions; Panel \textbf{B.}: correlations for the
    \textit{Birth of Stars} lecture and associated questions). The colors denote
    question identities. The diamonds in each panel denote the moment of peak
    correlation between the indicated question and the lecture trajectory. The
    associated questions' text and snippets of the lectures' transcripts from the surrounding 30 seconds, are displayed at the bottom of the
    figure.}

@@

    \caption{\textbf{Mapping out the geometry of knowledge and learning.}
    \textbf{A. Average ``knowledge maps'' estimated using each quiz.} Each map
    displays a 2D projection of the estimated knowledge about the content
    reflected by \textit{all} regions of topic space (see
    \nameref{subsec:knowledge-maps}). The topic trajectories of each
    lecture and the coordinates of each question are indicated by dotted lines
    and dots. Each map reflects an average across all participants. For
    individual participants' maps, see
    Figures~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,
    and~\individualKnowledgeMapsC. \textbf{B. Average ``learning maps''
    estimated between each successive pair of quizzes.} The learning maps are
    in the same general format as the knowledge maps in Panel A, but each
    coordinate in the learning maps indicates the \textit{difference} between
    the corresponding coordinates in the indicated \textit{pair} of knowledge
    maps---i.e., how much the estimated knowledge ``changed'' across the two
    quizzes. Each map reflects an average across all participants. For
    individual participants' maps, see
    Figures~\individualLearningMapsA~and~\individualLearningMapsB. \textbf{C.
    Word clouds for sampled points in topic space.} Each word cloud displays
    the relative weights of each word (via their relative sizes) reflected by the blend of topics
    represented at the locations of the stars in the maps. The words' colors
    indicate how much each word is weighted, on average, across all timepoints'
    topic vectors in the \textit{Four Fundamental Forces} (blue) and
    \textit{Birth of Stars} (green) videos, respectively.}

@@

We projected the resulting 100-dimensional topic vectors (for each second of
video and for each question) into a shared 2-dimensional space (see
\nameref{subsec:knowledge-maps}). Next, we sampled
points evenly from a $100 \times 100$ grid of coordinates that evenly tiled a
rectangle enclosing the 2D projections of the videos and questions. We used
Equation~\ref{eqn:rbf-knowledge} to estimate participants' knowledge at each of
these 10,000 sampled locations, and we averaged these estimates across
participants to obtain an estimated average \textit{knowledge map}
(Fig.~\ref{fig:knowledge-maps}). Intuitively, the knowledge map constructed
from a given quiz's responses provides a visualization of how ``much''
participants know about any content expressible by the fitted text embedding
model.

@@

Because the 2D projection we used to construct the knowledge and learning maps
is invertible, we may gain additional insights into the estimates
by reconstructing the original high-dimensional topic vectors for any point(s)
in the maps we are interested in. For example, this could serve as a useful
tool for an instructor looking to better understand which content areas a
student (or a group of students) knows well (or poorly). As a demonstration, we
show the top-weighted words from the blends of topics reconstructed from three
example locations on the maps (Fig.~\ref{fig:knowledge-maps}C): one point near
the \textit{Four Fundamental Forces} embedding (yellow); a second point near
the \textit{Birth of Stars} embedding (orange), and a third point somewhere in
between the two lectures' embeddings (pink). As shown in the word clouds in the
Panel, the top-weighted words at the example coordinate near the \textit{Four
Fundamental Forces} embedding also tended to be weighted heavily by the topics
expressed in that lecture. Similarly, the top-weighted words at the example
coordinate near the \textit{Birth of Stars} embedding tended to be weighted
most heavily by the topics expressed in \textit{that} lecture. And the
top-weighted words at the example coordinate between the two lectures'
embeddings show a roughly even mix of words most strongly associated with each
lecture.

@@

Over the past several years, the global pandemic has forced many educators to
teach remotely~\citep{MoseEtal21, ShimLee20, KawaEtal21, Whal20}. This change
in world circumstances is happening alongside (and perhaps accelerating)
geometric growth in the availability of high quality online courses on
platforms such as Khan Academy~\citep{Khan04}, Coursera~\citep{Youn12},
EdX~\citep{Kolo13}, and others~\citep{RhoaEtal13}. Continued expansion of the
global internet backbone and improvements in computing hardware have also
facilitated improvements in video streaming, enabling videos to be easily
downloaded and shared by large segments of the world's population. This
exciting time for online course instruction provides an opportunity to
re-evaluate how we, as a global community, educate ourselves and each other.
For example, we can ask: what makes an effective course or training program?
Which aspects of teaching might be optimized or automated? How can we provide  How and why do
learning needs and goals vary across people? How might we lower barriers to
achieving a high quality education?

@@

\subsubsection*{Constructing text embeddings of multiple lectures and questions}\label{subsec:topic-modeling}
We adapted an approach we developed in prior work~\citep{HeusEtal21} to embed
each moment of the two lectures and each question in our pool in a common representational space.
Briefly, our approach uses a topic model (Latent Dirichlet Allocation;~\citealp{BleiEtal03}), trained on a set
of documents, to discover a set of $k$ ``topics'' or ``themes.'' Formally, each
topic is defined as a set of weights over each word in the model's vocabulary
(i.e., the union of all unique words, across all documents, excluding ``stop
words.''). Conceptually, each topic is intended to give larger weights to words
that are conceptually related or that tend to co-occur in the same documents.
After fitting a topic model, each document in the training set, or any
\textit{new} document that contains at least some of the words in the model's
vocabulary, may be represented as a $k$-dimensional vector describing how much
the document (most probably) reflects each topic. (Unless, otherwise noted, we
used $k = 15$ topics.)

As illustrated in Figure~\ref{fig:sliding-windows}A, we start by building up a
corpus of documents using overlapping sliding windows that span each video's
transcript. Khan Academy provides professionally created, manual 
transcriptions of all videos for closed captioning. However, such transcripts 
would not be readily available in all contexts to which our framework could 
potentially be applied. Khan Academy videos are hosted on the YouTube 
platform, which additionally provides automated captions. We opted to use 
these automated transcripts (which, in prior work, we have found are 
sufficiently near-human quality yield reliable data in behavioral studies; 
\citealp{ZimaEtal18}) when developing our framework in order to make it 
more easily extensible and adaptable by others in the future.

We fetched these automated transcripts using the \texttt{youtube-transcript-api} 
Python package (\textcolor{red}{\textbf{Jeremy can you add a citation for this?} 
https://github.com/jdepoix/youtube-transcript-api}). The transcripts consisted of 
one timestamped line of text for every few seconds (mean: 2.34s; standard 
deviation: 0.83s) of spoken content in the video (i.e., corresponding to each 
individual caption that would appear on-screen if viewing the lecture via YouTube, 
and when those lines would appear). We
defined a sliding window length of (up to) $w = 30$ transcript lines, and
assigned each window a timestamp corresponding to the midpoint between its first
and last lines' timestamps. These sliding windows ramped up and down in length
at the very beginning and end of the transcript, respectively. In other words,
the first sliding window covered only the first line from the transcript; the
second sliding window covered the first two lines; and so on. This insured that
each line of the transcript appeared in the same number ($w$) of sliding
windows. After performing various standard text preprocessing (e.g., normalizing case, 
lemmatizing, removing punctuation and stop-words), we treated the text from each 
sliding window as a single ``document,''
and we combined these documents across the two videos' windows to create a single
training corpus for the topic model.  The top words from each of the 15 discovered
topics may be found in Table~\topics.

After fitting a topic model to each videos' transcripts, we could use the
trained model to transform arbitrary (potentially new) documents into
$k$-dimensional topic vectors. A convenient property of these topic vectors is
that documents that reflect similar blends of topics (i.e., documents that
reflect similar themes, according to the model) will yield similar (in terms of
Euclidean \textcolor{red}{should we say something like cosine instead of euclidean?} distance, correlation, etc.) topic vectors. In general, the
similarity between different documents' topic vectors may be used to
characterize the similarity in conceptual content between the documents.

We transformed each sliding window's text into a topic vector, and then used
linear interpolation (independently for each topic dimension) to resample the
resulting timeseries to one vector per second. We also used the fitted model to obtain topic
vectors for each question in our pool (Tab.~\questions). Taken together, we
obtained a \textit{trajectory} for each video, describing its path through
topic space, and a single coordinate for each question
(Fig.~\ref{fig:sliding-windows}C). Embedding both videos and all of the
questions using a common model enables us to compare the content from different
moments of videos, compare the content across videos, and estimate potential
associations between specific questions and specific moments of video.

@@

\begin{equation}
    \mathrm{ncorr}(x, y) = \frac{\mathrm{corr}(x, y) - \mathrm{mincorr}}{\mathrm{maxcorr} - \mathrm{mincorr}},
\end{equation}
and where $\mathrm{mincorr}$ and $\mathrm{maxcorr}$ are the minimum and maximum
correlations between any lecture timepoint and question, taken over all
timepoints and questions across both lectures and all five question used to estimate the knowledge trace \textcolor{red}{\textbf{Note: make sure this is correct in results section; not sure i caught all instances}}. We
also define $f(s, \Omega)$ as the $s$\textsuperscript{th} topic vector from the
set of topic vectors $\Omega$. Here $t$ indexes the set of lecture topic
vectors, $L$, and $i$ and $j$ index the topic vectors of questions used to estimate the knowledge trace, $Q$. Note that ``correct'' denotes the set of indices of
the questions the participant answered correctly on the given quiz.

Intuitively, $\mathrm{ncorr}(x, y)$ is the correlation between two topic
vectors (e.g., the topic vector from one timepoint in a lecture, $x$, and the
topic vector for one question, $y$), normalized by the minimum and maximum
correlations (across all timepoints $t$ and questions $Q$) to range between 0 and 1,
inclusive. Equation~\ref{eqn:prop} then computes the weighted average
proportion of correctly answered questions about the content presented at
timepoint $t$, where the weights are given by the normalized correlations
between timepoint $t$'s topic vector and the topic vectors for each question.
The normalization step (i.e., using $\mathrm{ncorr}$ instead of the raw
correlations) insures that every question
contributes some non-zero amount to the knowledge estimate.

@@

An important feature of our approach is that, given a trained text embedding
model and participants' quiz performance on each question, we can estimate
their knowledge about \textit{any} content expressible by the embedding 
model---not solely the content explicitly probed by the quiz questions or even 
appearing in the lectures. To visualize these estimates
(Figs.~\ref{fig:knowledge-maps},~\individualKnowledgeMapsA,~\individualKnowledgeMapsB,~\individualKnowledgeMapsC,~\individualLearningMapsA,
and~\individualLearningMapsB), we used Uniform Manifold Approximation and Projection (UMAP; \citealp{McInEtal18a}) to construct a 2D
projection of the text embedding space. Sampling the original 100-dimensional
space at high resolution to obtain an adequate set of topic vectors spanning
the embedding space would be computationally intractable. However, sampling a
2D grid is much more feasible. We defined a rectangle enclosing the 2D
projections of the lectures' and quizzes' embeddings, and we sampled points
from a regular $100 \times 100$ grid of coordinates that evenly tiled the
enclosing rectangle. We sought to estimate participants' knowledge (and
learning, i.e., changes in knowledge) at each of the resulting 10,000
coordinates.
